---
title: 'Race in the Oscars: Exploratory Analysis'
author: "Clarabelle McKeirnan"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

The Oscars Ceremonies have been repeatedly accused of racial bias in recent decades. The hashtag #OscarsSoWhite has been used countless times to address the issue of racial diversity, and whether or not the Oscars discriminates against People of Color. To explore this issue, we decided to select a dataset called "Race in the Oscars", which lists the winners and nominees for 92 different Oscar categories from 1928 to 2020. The dataset also lists the name, gender, and race of all the nominees and winners, with their respective film titles. In analyzing this dataset, we sought to find patterns, trends, and answers to the question  "Do the Oscars Ceremonies reflect racial diversity, or are the Oscars truly #SoWhite?

## Summary of Dataset

Because we wanted to analyze the racial diversity of the Oscars, we chose to focus on select rows and columns from the dataset. The size of the dataset originally contained over 10.4K rows and 9 columns. We narrowed this dataset down to focus solely on the years 1980 through 2020 and the racial diversity of the winners, by year and by category.


Our modified dataset contains 3,851 rows and 4 columns, and lists 92 different categories, such as Production Design, Cinematography, and Directing. There are four races listed: White, Asian, Hispanic, and Black. The dataset also lists whether or not the person who was nominated won, as well as the year of the ceremony. Because we wanted to focus on the racial diversity of the winners by year and category, we only included these respective columns.

```{r table_script,echo=FALSE,message=FALSE,tidy=TRUE}
library(dplyr)
library(knitr)
library(rmarkdown)
oscars<-read.csv(url("https://raw.githubusercontent.com/info201b-2022-spring/final-projects-swizzlejuice/main/data/oscars.csv?token=GHSAT0AAAAAABUUS75RCIQGQTGD5VLVKORMYUFKUOA"))
year_table <- oscars[-c(1:6544,10396), ]
table_script = year_table %>% group_by(year_ceremony)  %>%
  select(Category,
            Race,
            winner)
paged_table(table_script)
```

## Including Plots


```{r pressure, echo=FALSE}
plot(pressure)
```

